
\PassOptionsToPackage{xetex}{xcolor}
\PassOptionsToPackage{xetex}{graphicx}
\documentclass[a4paper,landscape,headrule,footrule,xetex]{foils}

\input{headx.tex}

\begin{document}
\header{Lecture 13}{Review and Feedback}{}
\maketitle

%\include{schedule}
\section{Where do we go from here?}
%\myslide{Parting Words}
\begin{quotation}
And so, reader, farewell to Sherlock Holmes! I thank you for your past constancy, and can but hope that some return has been made in the shape of that distraction from the worries of life and stimulating change of thought which can only be found in the fairy kingdom of romance.
\\ \mbox{} \hfill Arthur Conan Doyle (Preface to \href{http://compling.hss.ntu.edu.sg/canon/case.html}{The Case Book of Sherlock Holmes})
\end{quotation}


\myslide{Parting Words}

\begin{quotation}
  Education never ends, Watson. It is a series of lessons with the
  greatest for the last. This is an instructive case. There is neither
  money nor credit in it, and yet one would wish to tidy it up. When
  dusk comes we should find ourselves one stage advanced in our
  investigation. 
\\ \mbox{} \hfill  Sherlock Holmes
\href{http://compling.hss.ntu.edu.sg/canon/redc.html}{The Adventure of the Red Circle} 
\end{quotation}


\begin{quotation}
  Here I had heard what he had heard, I had seen what he had seen,
  and yet from his words it was evident that he saw clearly not only
  what had happened, but what was about to happen, while to me the
  whole business was still confused and grotesque. 
\\ \mbox{} \hfill Dr John Watson, \href{http://compling.hss.ntu.edu.sg/canon/redh.html}{The Red-Headed League}
\end{quotation}


\myslide{Reflection}
\MyLogo{Please email me your thoughts 
  \href{mailto:bond@ieee.org}{<bond@ieee.org>}}
\begin{itemize}
\item What was the most surprising thing in this class?
\item What do you think is most likely wrong?
\item What do you think is the coolest thing you've learned?
\item What do you think you're most likely to
remember?
\item How do you think this course will influence you from now on?
\item What (if anything) did you hope to learn that you didn't?
\end{itemize}
\bigskip\bigskip
\begin{center}
  This is the third time I have given this class \\
  I still want to make it better!
\end{center}







\end{document}

%%% Local Variables: 
%%% coding: utf-8
%%% mode: latex
%%% TeX-PDF-mode: t
%%% TeX-engine: xetex
%%% End: 

