\documentclass{beamer}
\input{semantics.tex}
\title[Syntax and Semantics]{Introduction}

\usepackage{forest}


\begin{document}

\maketitle

%\include{schedule}

\section{Welcome}
\begin{frame}{Welcome!}
\begin{itemize}
\item In this course we will introduce you to the study of grammar and meaning
  \begin{itemize}
  \item Syntax
    \begin{itemize}
    \item How words differ from each other
    \item How they can be combined to make larger structures
    \end{itemize}
  \item Semantics and Pragmatics
    \begin{itemize}
    \item How meaning is built up from words and phrases
    \item How meaning depends on context
    \end{itemize}
  \end{itemize}
\item We will ground the analysis with real examples from the Sherlock Holmes stories and the War with the Newts
  \begin{itemize}
  \item I try to make this as enjoyable as possible
  \end{itemize}
\end{itemize}

\end{frame}


\begin{frame}{Overview}
  \begin{itemize}
  \item The course is co-taught by Joanna Ut-Seong Sio (Syntax)
  \item The timetable is available in Moodle.   
  \item Most weeks you will have a lecture and a tutorial
   \\ (no tutorial on the first and last weeks)
  \item Assessment
    \begin{itemize}
    \item Satisfactorily complete ALL tutorial sets (40 marks) and attend 80\% of the tutorials. Students will discuss and complete the problem sets with the tutors during tutorials and hand them in at the next tutorial. 
    \item Do the assigned readings to make sure you understand the content of the course
    \item Two quizzes (60 marks, 30 marks each). Students must obtain 20 marks out of 30 marks for each test to pass the course. 
    \item People doing 5UJ2 get an extra small project (TBA).
    \end{itemize}
  \end{itemize}
\end{frame}


\begin{frame}{Textbook and Readings}

  \begin{itemize}
  \item Syntax: Carnie (2012) \textit{Syntax A Generative Introduction}, 3rd Edition. Wiley-Blackwell.
  \item Semantics: Saeed, John (2015) \textit{Semantics}. 4rd
  Edition. Wiley-Blackwell.
\end{itemize}

\begin{itemize}
\item If you want to know more about semantics I recommend
  \begin{itemize}
  \item Paul Kroeger (2022) \textit{Analyzing meaning: An introduction to semantics and pragmatics}. 3rd edition.  Language Science Press. \href{https://langsci-press.org/catalog/book/359}{DOI: 10.5281}  (Open Source)
%  \item Saeed, John (2015) \textit{Semantics}. 4rd Edition. Wiley-Blackwell. 
  \item Lyons, John (1977) \textit{Semantics}.  Cambridge University Press
  \end{itemize}
\end{itemize}
\end{frame}

\input{part/what-is-semantics}
%\input{part/self-intro}
\input{part/saw-a-kid} 
\input{part/referential-referring}
%\input{part/deixis}
\input{part/administrivia}








% \begin{frame}[allowframebreaks]
%   \frametitle{References}
%   \bibliographystyle{aclnat}
%   \bibliography{abb,mtg,ling,nlp}
% \end{frame}

\input{part/czech-terms-1.tex}

\end{document}

%%% Local Variables: 
%%% coding: utf-8
%%% mode: latex
%%% TeX-PDF-mode: t
%%% TeX-engine: xetex
%%% End: 

