\section{Administrivia}



\begin{frame}{Administrivia}
\begin{description}
\item [Coordinator]  Francis \ul{Bond} 
  {\small \url{<bond@ieee.org,francis.bond@upol.cz>}}
\item Course details are all online
\end{description}


\end{frame}

\begin{frame}{Extra Credit}

\begin{itemize}
\item If you submit a correction that gets accepted for one of the
  resources we use then it shows good mastery of the material
  \begin{itemize}
  \item you can get 1-5\% extra credit (depending on the size/difficulty)
\\ Mark $n \propto 10^{n-1}$ lines of code/documentation
  \item You can't go over 100\%
  \end{itemize}
\item A correction can involve
  \begin{itemize}
  \item fixing an error in transcription or annotation
    \begin{itemize}
    \item spelling error
    \item wrong sense
    \item error in the dictionary
    \end{itemize}
  \item making the documentation easy to read
  \item pointing out an error in a translation / finding a new translation
  \item coming up with a good example of a phenomenon
%  \item  fixing a bug in code
%  \item extending the code with new capabilities
  \end{itemize}

\end{itemize}



\end{frame}

\begin{frame}{Student Responsibilities}

By remaining in this class, the student agrees to:
\begin{enumerate}
\item  Make a genuine effort to learn and engage.
\item Read messages and participate.
\item Do assignments on time.
\item Attend regularly.
\item Seek help early, not last minute.
\item Treat peers respectfully.
\end{enumerate}

\end{frame}

\begin{frame}{Attendance}
\begin{enumerate}
\item You are expected to attend all classes.
\item Be on time - lateness is disruptive to your own and others' learning.
\item Valid reasons for missing class include the following:
\begin{enumerate}
\item A medical emergency (including mental health emergencies)
\item A family emergency (death, birth, natural disaster, etc).
\end{enumerate}
\item There will be significant material covered in class that is not in your readings.  You cannot expect to do well without coming to class.
\item If you miss a class, it is your responsibility to get the notes, any handouts you missed, schedule changes, etc. from a classmate.
\end{enumerate}

\end{frame}

\begin{frame}{Remediation and Academic Integrity}
\begin{enumerate}
\item No late work will be accepted, except in the case of a documented excuse.
\item For planned, justified, absences on class days or days on which assignments are due, advance notice must be provided.
\item Cheating will not be tolerated. Violations, including plagiarism, will be seriously dealt with, and could result in \textbf{a failing grade for the entire course}.
\item Refer to the University Honour Code
\item As always, use your common sense and conscience.
\end{enumerate}


\end{frame}


\begin{frame}{The winning strategy}

\begin{itemize}
\item Read the textbook before class (and after again, if necessary)
\item Work together: make study groups
\item Tasks: Discuss as much as you want, annotate your own answers
\item Tutorials: Discuss as much as you want, write your own answers
\item Treat LLMs as though they were a human --- you can discuss the
  answer with them, but they should not give you the entire answer
  
\item Ask questions \ldots\ early and often!
\end{itemize}

\end{frame}
